---
title: "Data 607 Assignment 3"
author: "Carol Campbell"
date: "2023-09-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r nclude=FALSE}
library(tidyverse)
library(stringr)
library(dplyr)

```

#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
#use read.csv() to load data from github

col_majors <- read.csv ("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv", header = TRUE, sep = ",")

#filter for Data or Statistics like majors  
col_major_mod <- col_majors %>%
  filter(str_detect(Major, "DATA|STATISTICS"))

glimpse(col_major_mod)

```

#2 Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
The two exercises below are taken from R for Data Science:

```{r}

produce <- data.frame(c("bell pepper", "bilberry", "blackberry","blood orange","blueberry","cantalope","chili pepper","cloudberry","elderberry","lime","lychee","mulberry","olive","salal berry"))

#view produce
produce

#paste produce collapsed by , into concatenate
cat(paste0(produce), collapse=",")
```

#3 Describe, in words, what these expressions will match:
(.)\1\1 
ANS - "(.)\1\1" looks for any character that is followed by "\1\1"

"(.)(.)\\2\\1" 
ANS - "(.)(.)\\2\\1" looks for any two characters that are repeated in reverse, ie ace, eca

(..)\1 
ANS - (..)\1 finds any word with characters followed by  "\1"

"(.).\\1.\\1" 
ANS - "(.).\\1.\\1" searches for any character, "a" followed by any other character, followed by original character "a", followed by any other character, and the original character again

"(.)(.)(.).*\\3\\2\\1" - 
ANS - This expression searches for any three characters followed by same three characters in reverse order

```{r}
# (.)\1\1 
my_string <- (c("abracadabra", "emme","farm\1\1", "aaa","civic", "peep", "aabbaa", "ccc" ))

# find any where character is followed by "\1\1"
x <- str_view(my_string, "(.)\1\1")   
print(x)

# find any two characters where they are repeated in reverse, ie ab, ba
y <- str_view(my_string, "(.)(.)\\2\\1")
print(y)

# find any where word with characters followed by literal "\1"
z <- str_view(my_string, "(..)\1")
print(z)

a <- str_view(my_string, "(.).\\1.\\1")
print(a)

b <- str_view(my_string, "(.)(.)(.).*\\3\\2\\1")
print(b)

```

#4 Construct regular expressions to match words that:
Start and end with the same character.
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
#expression to find words that start and end with the same character
mystring2 <- c("rubber", "Emme", "Toronto", "decide", "melee", "termite" )
sc <- str_subset(mystring2, regex("^(.).*\\1$", ignore_case = T))
print (sc)

#expression that contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
rc <- str_subset(mystring2, regex("([A-Za-z][A-Za-z]).*\\1", ignore_case = T))
print (rc)

#one letter repeated in at least three places
tr <- str_subset(mystring2, regex("([A-Za-z]).*\\1.*\\1", ignore_case = T))
print (tr)
```

