---
title: "Data 607 Project 1 Chess Scores2"
author: "Carol Campbell"
date: "2023-10-01"
output: html_document
---

In this project, we’re given a text file with chess tournament results where the information has some structure. Our job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

For the first player, the information would be:

Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load libraries used
```{r message=FALSE}
library(tidyverse)
library(stringr)
library(ggplot2)
library(readr)
library(kableExtra)
```

##Read file from Github
```{r}
chess_tbl = readLines("https://raw.githubusercontent.com/carolc57/Data607-Fall23/main/chess_scores.txt")
head(chess_tbl, 15)

```

##Remove rows w/ dashes
```{r}
#remove rows w/ dashes
chess_tbl_mod <- str_replace_all(string = chess_tbl, pattern = "^-+$", "")
head (chess_tbl_mod)
```
#Remove empty vectors
```{r}
chess_tbl_mod <- chess_tbl[sapply(chess_tbl_mod, nchar) > 0]
head(chess_tbl_mod)
```
##Remove Header rows
```{r}
chess_tbl_mod <- chess_tbl_mod[-c(1:2)]  ##remove rows 1 and 2 because interfere w/ later operations
head(chess_tbl_mod, 16)

```

##Since the data for one competitor is spread between two rows, I’ll need to create two separate dataframe and paste them together 
```{r}
chess_tbl_mod_odd = chess_tbl_mod[seq(1,128,2)]
chess_tbl_mod_even = chess_tbl_mod[seq(2,128,2)]
```

#The combined data frame
```{r}
chess_tbl_mod_combined <- paste(chess_tbl_mod_odd, chess_tbl_mod_even)
head(chess_tbl_mod_combined, 16)

```
#Tidy Player data. Use various str and regex functions to parse data
```{r}
chess_final <- chess_tbl_mod_combined  #change file name to preserve data prior to manipulation
head(chess_final)


PlayerIDTemp <- as.integer(str_extract(chess_final, "\\d+"))
PlayerID <- subset(c(PlayerIDTemp), c(PlayerIDTemp)!="0",64)


Name <- str_extract(string = chess_final, pattern = "\\s([[:alpha:] ]{5,})\\b\\s")  #gets name

Player_Points <- str_extract(string = chess_final, pattern = "[0-9]\\.[0-9]")  #gets total points for each player

State <- unlist(str_extract_all(chess_final, "\\|[[:space:]]{1,}[[A-Z]]{2} \\|"))
State <- str_replace_all(State, pattern = "(\\|[[:space:]]{1,})|([[:space:]]{1,}\\|)", replacement = "")

Player_rating <- str_extract(string = chess_final, pattern = "\\s\\d{3,4}[^\\d]")
#remove "P" at end of rating for some players
Player_rating <- as.integer(str_extract(Player_rating, "\\d+"))

USCF_id <- str_extract(string = chess_final, pattern ="[[:digit:]]{8}") 

```

#Opponent Data - parse player's opponets per round
```{r}
Opp_Id <-str_extract_all(str_extract_all(chess_final, "\\d+\\|"), "\\d+")
Opp_Id <-subset(c(Opp_Id), c(Opp_Id)!="0")
head(Opp_Id)
```
##Function to calculate opponent average ratings
```{r}
x<-length(chess_tbl_mod)
OppAvgRtg <-numeric(x/2)
for (i in 1:(x/2)) {
  OppAvgRtg[i] <- mean(Player_rating[as.numeric(unlist(Opp_Id[PlayerID[i]]))])
}
OppAvgRtg <- round((OppAvgRtg),0)
```

##chess_ratings dataframe
```{r}
chess_ratings <- data.frame(Name,State,Player_Points,Player_rating, OppAvgRtg)

```

##Write output to .csv file
```{r}
write.csv(chess_ratings, file = "Carols Chess Ratings.csv");

```

##Chess Ratings Table
```{r}
kable(chess_ratings) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "center")
```