---
title: "Data 607 Sentiment Analysis Part 1"
author: "Carol Campbell"
date: "2023-11-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```

###  Part 1 of 2 - recreate the code from Chapter 2 of the book Text Mining with R by Juilia Silge and David Robinson. https://www.tidytextmining.com/sentiment.html

## **CHAPTER 2 SENTIMENT ANALYSIS WITH TIDY DATA**

## Packages used for this assignment
```{r}
#install.packages("tidyverse")
#install.packages("textdata")
#install.packages("gutenbergr")
#install.packages("DT")
#install.packages("flextable")
#install.packages("wordcloud")
```



## Install libaries
```{r}

library(tidyverse)
library(tidyr)
library(ggplot2)
library(wordcloud)
library(reshape2)

```

## **THE SENTIMENTS DATASETS**

### Afinn lexicon assigns words with a score between -5 and 5,  indicating negative sentiment or positive sentiment, and rating thereof.

```{r}
# Load Libraries

library(tidyverse)
library(tidytext)
library(textdata)

# load afinn lexicons 

get_sentiments("afinn")

```

### Use bing lexicon to categorize words in a binary fashion of positive and negative categories.

```{r}
get_sentiments("bing")
```

###  The nrc lexicon categorizes words in a binary fashion (“yes”/“no”) into categories of positive, negative, anger, anticipation, disgust, fear, joy, sadness, surprise, and trust.

```{r}
get_sentiments("nrc")
```
## **SENTIMENT ANALYSIS WITH INNER JOIN**

### Take a look words with a joy score from the NRC lexicon. To do so, we take the text of the novels and it (the text) to the tidy format using unnest_tokens(). Create other columns to keep track of which line and chapter of the book each word comes from, using group_by and mutate to construct those columns.

```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() |>
  group_by(book) |>
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) |>
  ungroup() |>
  unnest_tokens(word, text)
```

### Sentiment analysis can begin now that the text is in a tidy format of one word per row. First, let’s use the NRC lexicon and filter() for the joy words. Next, let’s filter() the data frame with the text from the books for the words from Emma and then use inner_join() to perform the sentiment analysis. What are the most common joy words in Emma? Let’s use count() from dplyr.

### There are mostly positive, happy words about hope, friendship, and love here. We also see some words that may not be used joyfully by Austen (“found”, “present”)

```{r}
nrc_joy <- get_sentiments("nrc") |> 
  filter(sentiment == "joy")

tidy_books |>
  filter(book == "Emma") |>
  inner_join(nrc_joy) |>
  count(word, sort = TRUE)
```
### Here we examine how sentiment changes throughout each novel. We can do this using a few dplyr functions. First, we find a sentiment score for each word using the Bing lexicon and inner_join().

### Next count up how many positive and negative words there are in defined sections of each book. We define an index here to keep track of where we are in the narrative; this index (using integer division) counts up sections of 80 lines of text.

### The %/% operator does integer division (x %/% y is equivalent to floor(x/y)) so the index keeps track of which 80-line section of text we are counting up negative and positive sentiment in.

### Use pivot_wider() so that negative and positive sentiment appear in separate columns, and lastly calculate the net sentiment (positive - negative).

```{r}

jane_austen_sentiment <- tidy_books |>
  inner_join(get_sentiments("bing")) |>
  count(book, index = linenumber %/% 80, sentiment) |>
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |> 
  mutate(sentiment = positive - negative)

```
## We sentiment through the narratives of Jane Austen’s novels

### Plot these sentiment scores across the plot trajectory of each novel. We are plotting against the index on the x-axis that keeps track of narrative time in sections of text.

### Here we see how the plot of each novel changes toward more positive or negative sentiment over the trajectory of the story.

```{r}

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")

```
## **COMPARING THE THREE SENTIMENT DICTIONARIES**

### In order to determine which sentiment lexicon is appropriate for our analysis, we will use all three sentiment lexicons and examine how the sentiment changes across the narrative arc of Pride and Prejudice. First, we use filter() to choose only the words from the one novel we are interested in.

```{r}
pride_prejudice <- tidy_books |> 
  filter(book == "Pride & Prejudice")

pride_prejudice
```

## Use inner_join() to calculate the sentiment in different ways.

## To find a sentiment score in chunks of text throughout the novel, we need to use a different pattern for the AFINN lexicon than we did for
the other two.

### Use integer division (%/%) to define larger sections of text that span multiple lines, and we can use the same pattern with count(), pivot_wider(), and mutate() to find the net sentiment in each of these sections of text.

```{r}
afinn <- pride_prejudice |> 
  inner_join(get_sentiments("afinn")) |> 
  group_by(index = linenumber %/% 80) |> 
  summarise(sentiment = sum(value)) |> 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice |> 
    inner_join(get_sentiments("bing")) |>
    mutate(method = "Bing et al."),
  pride_prejudice |> 
    inner_join(get_sentiments("nrc") |> 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) |>
    mutate(method = "NRC")) |>
  count(method, index = linenumber %/% 80, sentiment) |>
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) |> 
  mutate(sentiment = positive - negative)
```

## Compare the three sentiment lexicons using Pride and Prejudice

## Now that we have an estimate of the net sentiment (positive - negative) in each chunk of the novel text for each sentiment lexicon, 
bind them together and visualize them.

## The three different lexicons for calculating sentiment give results that are different in an absolute sense but have similar relative trajectories through the novel. There are similar dips and peaks in sentiment at about the same places in the novel, but the absolute values are vastly different. The AFINN lexicon has the largest absolute values, with high positive values. The lexicon from Bing et al. has lower absolute values and seems to label larger blocks of contiguous positive or negative text. The NRC results are shifted higher relative to the other two, labeling the text more positively, but detects similar relative changes in the text. We find similar differences between the methods when looking at other novels; the NRC sentiment is high, the AFINN sentiment has more variance, the Bing et al. sentiment appears to find longer stretches of similar text, but all three agree roughly on the overall trends in the sentiment through a narrative arc.

```{r}
bind_rows(afinn, 
          bing_and_nrc) |>
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

## Notice how the result for the NRC lexicon biased  high in sentiment compared to the Bing et al. result. One can see how many positive and negative words are in these lexicons.

## Both lexicons have more negative than positive words, but the ratio of negative to positive words is higher in the Bing lexicon than the NRC lexicon. This will contribute to the effect we see in the plot above, as will any systematic difference in word matches. Whatever the source of these differences, note the similar relative trajectories across the narrative arc, with similar changes in slope, but marked differences in absolute sentiment from lexicon to lexicon. This is an important consideration to keep in mind when choosing a sentiment lexicon for analysis.

```{r}
get_sentiments("nrc") |> 
  filter(sentiment %in% c("positive", "negative")) |> 
  count(sentiment)
```

```{r}
get_sentiments("bing") |> 
  count(sentiment)
```
### **MOST COMMON POSITIVE AND NEGATIVE WORDS**

## Here we analyze word counts that contribute to each sentiment. By using count()  with arguments of both word and sentiment, we identify how much each word contributed to each sentiment.

```{r}
bing_word_counts <- tidy_books |>
  inner_join(get_sentiments("bing")) |>
  count(word, sentiment, sort = TRUE) |>
  ungroup()
```
## Words that contribute to positive and negative sentiment in Jane Austen’s novels

## Pipe straight into ggplot2.

```{r}
bing_word_counts |>
  group_by(sentiment) |>
  slice_max(n, n = 10) |> 
  ungroup() |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```
## The word “miss” with a lower case "m" is coded as negative, but when typed with a capital "M", it is used as a title for young, unmarried women in Jane Austen’s works. One way to avoid this is to add “Miss” to a custom stop-words list using bind_rows(). 

```{r}
custom_stop_words <- bind_rows(tibble(word = c("Miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```
## **WORDCLOUDS**

###  The most common words in Jane Austen’s novels.

### A wordcloud, from the wordcloud package of base R graphics, allows us to see the most common words in Jane Austen’s works as a whole again.

```{r}
tidy_books |>
  anti_join(stop_words) |>
  count(word) |>
  with(wordcloud(word, n, max.words = 100))

```

### Most common positive and negative words in Jane Austen’s novels

### In order to use the function comparison.cloud(), you may need to turn the data frame into a matrix using reshape2’s acast(). We do the sentiment analysis to tag positive and negative words using an inner join, then find the most common positive and negative words. Until the step where we need to send the data to comparison.cloud().

### The size of a word’s text is proportionate to its frequency within its sentiment. The most important positive and negative words are revealed, but the sizes of the words are not comparable across sentiments.

```{r}

tidy_books |>
  inner_join(get_sentiments("bing")) |>
  count(word, sentiment, sort = TRUE) |>
  acast(word ~ sentiment, value.var = "n", fill = 0) |>
  comparison.cloud(colors = c("brown", "green"),
                   max.words = 100)
```
## **LOOKING AT UNITS BEYOND JUST WORDS**

### Some sentiment analysis algorithms look beyond only unigrams (i.e. single words) to try to understand the sentiment of a sentence as a whole. R packages included coreNLP (T. Arnold and Tilton 2016), cleanNLP (T. B. Arnold 2016), and sentimentr (Rinker 2017) are examples of such sentiment analysis algorithms.

```{r}
p_n_p_sentences <- tibble(text = prideprejudice) |> 
  unnest_tokens(sentence, text, token = "sentences")

```

```{r}
p_n_p_sentences$sentence[2]

```
### The sentence tokenizing seem to have a bit of trouble with UTF-8 encoded text, especially with sections of dialogue; it does much better with punctuation in ASCII. Before unnesting, we use the function unnest_tokens() to split into tokens using a regex pattern. We could use this to split the text of Jane Austen’s novels into a data frame by chapter.

### We have recovered the correct number of chapters in each novel (plus an “extra” row for each novel title). In the austen_chapters data frame, each row corresponds to one chapter.

```{r}
austen_chapters <- austen_books() |>
  group_by(book) |>
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") |>
  ungroup()

austen_chapters |> 
  group_by(book) |> 
  summarise(chapters = n())
```
### We use tidy text analysis to ask questions such as what are the most negative chapters in each of Jane Austen’s novels? 

### First, get the list of negative words from the Bing lexicon. 

### Second, make a data frame of how many words are in each chapter so we can normalize or the length of chapters. 

### Then, find the number of negative words in each chapter and divide by the total words in each chapter. 

### These are the chapters with the most sad words in each book, normalized for number of words in the chapter. 

#### Chapter 34 in Pride and Justice has the highest proportion of negative words with ratio of .053.

```{r}
bingnegative <- get_sentiments("bing") |> 
  filter(sentiment == "negative")

wordcounts <- tidy_books |>
  group_by(book, chapter) |>
  summarize(words = n())

tidy_books |>
  semi_join(bingnegative) |>
  group_by(book, chapter) |>
  summarize(negativewords = n()) |>
  left_join(wordcounts, by = c("book", "chapter")) |>
  mutate(ratio = negativewords/words) |>
  filter(chapter != 0) |>
  slice_max(ratio, n = 1) |> 
  ungroup()
```
##**SUMMARY**

### Sentiment analysis provides a way to understand the attitudes and opinions expressed in written word as intended by the author. To conduct a successful sentiment analysis use an inner join on tidy data structures. 