---
title: "Data 607-Homework-5"
author: "Carol Campbell"
date: "2023-10-07"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Assignment – Tidying and Transforming Data

We are given a .csv file containing disjointed data for two airlines - "Alaska" and "AM WEST", the five airports that they operate out of, and their respective arrival and departure delays.
Use 'tidyr' and 'dplyr' to tidy and transform the data.  Perform an analysis to compare the arrival delays for the two airlines. 


```{r load libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)
```

Read data from my github repository
```{r}
airlines <- read.csv(file = "https://raw.githubusercontent.com/carolc57/Data607-Fall23/main/Airlines.csv", header = TRUE, sep = ",", na.strings = "")

#display data
kable(airlines, format = "markdown")

#rename airlines to preserve original file
t_airlines <- airlines
```

Let the tidying begin...
```{r}
#rename columns 1 and 2 
names(t_airlines)[1] <- "Airlines"
names(t_airlines)[2] <- "Arrival_Status"

#Add airline names to blank column adjacent to delayed status 
t_airlines[2,1] <- "ALASKA"
t_airlines[5,1] <- "AM WEST"

#remove NA rows
t_airlines <- t_airlines[!apply(is.na(t_airlines[1:7]),1,all), ]

kable(t_airlines, format = "markdown")
```



Transform data from wide to long; filter for delayed flights only
```{r}
#transform from wide to long

t_airlines2 <- t_airlines  |>
  pivot_longer(cols = 3:7, 
               names_to = "City",
               values_to = "Count") |>
  filter(Arrival_Status =="delayed")            #because we're only interested in delayed flights for our analysis

kable(t_airlines2,format = "markdown")
```

Lets see the performance of these airlines by city
```{r}

flightsbycity <- t_airlines2 |> select(Airlines, City, Count) %>% 
      group_by(Airlines, City) %>% 
      summarise(total = sum(Count)) %>% 
      mutate(percent           = round(total/sum(total) * 100, 2),
             percent_formatted = paste0(round(total/sum(total) * 100, 1), "%"),
             total_flights     = sum(total)) 

flights_city <- as.data.frame(flightsbycity)

kable(flights_city, format = "markdown")
```


Graphical presentation of airline performance by city
```{r}
flights_city %>% 
ggplot( aes(x=City, y=percent, fill=Airlines)) +
    geom_bar(stat="identity", position=position_dodge(), colour="navy") +
    geom_text(aes(label=percent_formatted), vjust=.5, hjust=1,position= position_dodge(width=0.9),  color="navy") +
        ggtitle("Percentage of Flights Delayed by Airline by City") +
    xlab("Destination") + ylab("Percentage of Flights Delayed") +
        coord_flip() 
```
It's interesting to see that Alaska Airlines experienced the highest delays at Seattle while AM West had a higher number of delays at Phoenix. Both airlines experienced relatively close arrival delays at Los Angeles and San Francisco.   

Additional tidying to compare arrival delays between the two airlines
```{r}

stats_t_airlines2 <- t_airlines2 %>% dplyr::group_by(Airlines) %>% 
  dplyr::summarise(total=sum(Count),
                   max=max(Count), 
                   min=min(Count),
                   avg=mean(Count)
  )

stats_t_airlines2

airlines_comp <- as.data.frame(stats_t_airlines2)

kable(airlines_comp, format = "markdown")
```
From our analysis, we can conclude that Alaska Airlines has a better on-time performance than AM West Airlines over all cities.  Alaska Airlines average number of delayed flights was 100 compared to AM West Airlines' average number of delayed flights of 157.

Graphical representation of Airlines Delay
```{r}
airlines_comp |> 
    ggplot( aes(x=Airlines, y=total, fill=Airlines)) +
    geom_bar(stat="identity", position=position_dodge(), colour="navy", width = 0.5) +
    ggtitle("Flights Delayed by Airline") +
    xlab("Airlines") + ylab("Number of Flights Delayed") 
   
```
Again we can see clearly that Alaska Airlines had fewer delayed flights than AM West. 
